---
title: "Tetraploidy vs Diploidy evolutionary rescue comparison in the case of Heterozygot advantage"
authors: "Rüya Eylül Arslan and Aigerim Adilbekova"
contributions: "Prof. Dr. Claudia Bank, Dr. Loïc Marrec, Dr. Catalina Chaparro Pedraza, ChatGPT"
date: "2024-10-15"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

# Comperative modeling of Tetraploid and Diploid populations

To compare evolutionary rescue likelihood of two different ploidy
levels, we implemented two models: one for diploid populations and
another for tetraploid populations. These models simulate population
dynamics across generations.

To have a clear understanding, we imagined two different plant
populations with different ploidy under herbicide pressure. In both
models, we choose to set initial populations to consist of only
homozygous recessive(non-resistant) genotypes (for diploids "rr", for
tetraploids "rrrr") to simulate a simplified, controlled starting point,
before any resistancy evolved. Then by the mutations occuring, they will
acquire resistancy. By varying the selective advantage across genotypes,
we modeled a scenario where heterozygotes have an advantage over
homozygotes, reflecting real-world selective pressures in diverse
environments.

## Key similarities and distinctions between the models:

We used Hardy-Weinberg principles for deriving offspring genotypes for
both of the models. We used the generalization for the tetraploids in
our calculations. Tetraploids have a more complex inheritance by
structure.

Diploid Model: Incorporates two alleles per individual. Tetraploid
Model: Incorporates four alleles per individual. Tetraploid model is
expected to display increased genetic variation due to multiple allele
combinations.

We generated a table with varying selection coefficients to use later in
our analyses.

### list of Variables:

-   Diploid model

Number of rr individuals

Number of Rr individuals

Number of RR individuals

-   Tetraploid model

Number of rrrr individuals

Number of Rrrr individuals

Number of RRrr individuals

Number of RRRr individuals

Number of RRRR individuals

### list of Parameters:

-   Shared in both models

Decay rate(decay_rate) : It is the power of herbicide pressure to kill
individuals (\>0 to have potential extinction)

Selection coefficent(sel_coeffs): Selective advantage of the
heterozygote over other genotypes, its fitness. We generated a table
with varying levels of selection coefficients later. We have different
levels for the different genotypes we have. There is three different
sel_coeffs for the diploid model while five different for the tetraploid
model.

T_max = maximum generation number.

Mutation rate(mut_rate) : Must be bigger than 0 to evolve resistancy. We
used the same mutation rate for both the models, but to see if the
tetraploid does better only because of its ability to mutate double the
rate of the diploids, we later tested and compared the diploids with
double the mutation rate.

-   Diploid model

Initial population size of rr individuals (N_rr) : we set it to 250
initially Initial population size of Rr individuals (N_Rr) : we set it
to 0 initially Initial population size of RR individuals (N_RR) : we set
it to 0 initially

-   Tetraploid model

Initial population size of rrrr individuals (N_rrrr) : we set it to 250
initially Initial population size of Rrrr individuals (N_Rrrr) : we set
it to 0 initially Initial population size of RRrr individuals (N_RRrr) :
we set it to 0 initially Initial population size of RRRr individuals
(N_RRRr) : we set it to 0 initially Initial population size of RRRR
individuals (N_RRRR) : we set it to 0 initially

# simulating one generation in a diploid population

We want to observe the numbers of each diploid genotype (rr,Rr,RR) over
time. For this, we first need to simulate one generation at a time. At
generation $t$ the population consists of $N_rr$ homozygous wild-types,
$N_Rr$ heterozygous mutants and $N_RR$ homozygous mutants.

To simulate mutation from r to R, we model as a Poisson random variable
with a rate proportional to the mutation rate. The process of
reproduction is modeled by calculating the total gametes produced by
each genotype, followed by random mating according to the hardy weinberg
principles.

Table for expected frequencies for diploids:

p : the frequency of R

q : the frequency of r

| Genotype | Frequency |
|----------|-----------|
| RR       | p^2^      |
| Rr       | 2pq       |
| rr       | q^2^      |

Then we apply the selection coefficiencies of each genotype to calculate
the number of individuals in the next generation. For the diploid
population we take the 1st(for rr), 3rd(for Rr), 5th(for RR) values of
the selection coefficiency vector, to later use the same table we
generated for the tetraploid populations.

This method allows us to observe changes in genotype frequencies under
selective pressure and mutation over time, showing how the population’s
genetic structure evolves.

### function to simulate one generation

```{r}
# Function to simulate one generation for diploid genotypes
simulate_one_gen_diploid <- function(N_RR, N_Rr, N_rr, decay_rate, sel_coeffs, mut_rate) {
  
  #calculating the total population size
  total = N_RR + N_Rr + N_rr
  
  #calculating total gametes
  total_R = N_RR *2 + N_Rr
  total_r = N_rr *2 + N_Rr
  
  #mutating r to R
  mut_r_to_R <- rpois(1, total_r * mut_rate)
  
  #calculating gametes after mutation
  total_R_aftermut <- max(total_R + mut_r_to_R, 0)
  total_r_aftermut <- max(total_r - mut_r_to_R, 0)
  
  #calculating gamet frequencies p and q
  freq_p <- total_R_aftermut / (2 * total)
  freq_q <- total_r_aftermut / (2 * total)
  
  
  #hardy weinberg random mating
  freq_RR <- freq_p^2
  freq_Rr <- (2 * freq_p * freq_q)
  freq_rr <- freq_q^2
  
  #return to "expected" by multiplying freq with the pops
  N_RR <- freq_RR * total
  N_Rr <- freq_Rr * total
  N_rr <- freq_rr * total
  
  # Apply decay and selection coefficients for each genotype
  offsp_RR <- rpois(1, N_RR * (1 - decay_rate + sel_coeffs[1]))
  offsp_Rr <- rpois(1, N_Rr * (1 - decay_rate + sel_coeffs[3]))
  offsp_rr <- rpois(1, N_rr * (1 - decay_rate + sel_coeffs[5]))
  

  return(c(offsp_RR, offsp_Rr, offsp_rr))
}

#test the function
print(simulate_one_gen_diploid(N_RR=0, N_Rr=0, N_rr=100, 0.1, c(0,0.1,0.2,0.1,0), 0.01))
```

# simulating diploid population

```{r}
simulate_pop_diploid <- function(N_RR, N_Rr, N_rr, decay_rate, sel_coeffs, mut_rate,t_max) {
  # Create the vector in which to save the results
  pop_vector <- c(N_RR, N_Rr, N_rr)
  # initiate the variables
  pop_new <- c(N_RR, N_Rr, N_rr)
  
  #take the values of initial variables
  init_RR <- N_RR
  init_Rr <- N_Rr
  init_rr <- N_rr
  
  # run the simulation until generation t_max
  for (i in 1:t_max+1) {
    # redefine the current population one generation later
    pop_new <- simulate_one_gen_diploid(pop_new[1],pop_new[2],pop_new[3], decay_rate, sel_coeffs, mut_rate)
    # add the new population sizes to the output vector
    pop_vector <- rbind(pop_vector,pop_new)
    # condition to stop the simulation before t_max: either the population exceeds 1.5 times the original population size, or it goes extinct
    if (pop_new[1]+pop_new[2]+pop_new[3] >= 5*(init_RR + init_Rr + init_rr) | pop_new[1]+pop_new[2]+ pop_new[3]==0) break
  }
  
  # define the row and column names of the output vector
  rownames(pop_vector) <- (0:t_max)[1:length(pop_vector[,1])] # note that the vector has to be cut if the simulation stopped early
  colnames(pop_vector) <- c("RR","Rr","rr")
  # return the result
  return(pop_vector)	
}

output <- simulate_pop_diploid(N_RR = 0, N_Rr=0, N_rr=1000, decay_rate=0.175, sel_coeffs = c(0,0.174,0.25,0.174,0), mut_rate=0.05,t_max = 100)

# show the last few lines of the data table
print(output)

# plot the output - note that the plot range and the x axis have to be adjusted based on the length of the table
# determine x axis range 
x_range <- 0:(length(output[,1])-1)
# this plots the total population size
plot(x_range,output[,1]+output[,2]+output[,3],type='l',ylim=c(0,max(output[,1]+output[,2]+ output[,3])),xlab = "Generation",ylab = "Population size")
# add number of wild type individuals
lines(x_range,output[,1], col="blue")
# add number of mutant individuals
lines(x_range,output[,2], col="red")
lines(x_range,output[,3], col="green")
```

# Everything Tetraploid NOW!!!

# simulating one generation in a tetraploid population

We want to observe the numbers of each tetraploid genotype
(rrrr,Rrrr,RRrr,RRRr,RRRR) over time. For this, we first need to
simulate one generation at a time. At generation $t$ the population
consists of $N_rrrr$ homozygous wild-types, $N_Rrrr$ heterozygous
mutants, $N_RRrr$ fully-heterozygous mutants, $N_RRRr$ heterozygous
mutants, and $N_RRRR$ homozygous mutants.

To simulate mutation from r to R, we model as a Poisson random variable
with a rate proportional to the mutation rate. The process of
reproduction is modeled by calculating the total gametes produced by
each genotype, followed by random mating according to the Hardy-weinberg
principles generalization for polyploids.

Table for expected frequencies for tetraploids:

p : the frequency of R

q : the frequency of r

| Genotype | Frequency |
|----------|-----------|
| RRRR     | p^4^      |
| RRRr     | 4p^3^q    |
| RRrr     | 6p^2^q^2^ |
| Rrrr     | 4pq^3^    |
| rrrr     | q^4^      |

Then we apply the selection coefficiencies of each genotype to calculate
the number of individuals in the next generation. For the diploid
population we take the 1st(for rrrr), 2nd(for Rrrr), 3rd(for RRrr),
4th(RRRr), 5th(for RRRR) values of the selection coefficiency vector, to
later use the same table we generated for the diploid populations.

This method allows us to observe changes in genotype frequencies under
selective pressure and mutation over time, showing how the population’s
genetic structure evolves.

# simulating one generation function

```{r}
# Function to simulate one generation for tetraploid genotypes
simulate_one_gen_tetraploid <- function(N_RRRR_new, N_RRRr_new, N_RRrr_new, N_Rrrr_new, N_rrrr_new, decay_rate, sel_coeffs, mut_rate) {
  
    #computing total new population
  total_new <- N_RRRR_new + N_RRRr_new + N_RRrr_new + N_Rrrr_new + N_rrrr_new

    #calculate just the total gamete R and r 
  total_R <- (N_RRRR_new * 4 + N_RRRr_new * 3 + N_RRrr_new * 2 + N_Rrrr_new * 1)
  total_r <- (N_rrrr_new * 4 + N_Rrrr_new * 3 + N_RRrr_new * 2 + N_RRRr_new * 1)
  
  
  #and apply the mutation rate and then do the division
  mut_r_to_R <- rpois(1, total_r * mut_rate)
  
  total_R_aftermut <- max(total_R + mut_r_to_R, 0)
  total_r_aftermut <- max(total_r - mut_r_to_R, 0)
  
  freq_p <- total_R_aftermut / (4 * total_new)
  freq_q <- total_r_aftermut / (4 * total_new)
  
  #hardy weinberg random mating
  freq_RRRR <- (freq_p^4)
  freq_RRRr <- (4 * freq_p^3 * freq_q)
  freq_RRrr <- (6 * freq_p^2 * freq_q^2)
  freq_Rrrr <- (4 * freq_q^3 * freq_p)
  freq_rrrr <- (freq_q^4)
  
  #return to "expected" by multiplying freq with the pops
  N_RRRR_new <- freq_RRRR * total_new
  N_RRRr_new <- freq_RRRr * total_new
  N_RRrr_new <- freq_RRrr * total_new
  N_Rrrr_new <- freq_Rrrr * total_new
  N_rrrr_new <- freq_rrrr * total_new

  # Apply decay and selection coefficients for each genotype draw with rpois
  offsp_RRRR <- rpois(1, N_RRRR_new * (1 - decay_rate + sel_coeffs[1]))
  offsp_RRRr <- rpois(1, N_RRRr_new * (1 - decay_rate + sel_coeffs[2]))
  offsp_RRrr <- rpois(1, N_RRrr_new * (1 - decay_rate + sel_coeffs[3]))
  offsp_Rrrr <- rpois(1, N_Rrrr_new * (1 - decay_rate + sel_coeffs[4]))
  offsp_rrrr <- rpois(1, N_rrrr_new * (1 - decay_rate + sel_coeffs[5]))
  
return(c(offsp_RRRR, offsp_RRRr, offsp_RRrr, offsp_Rrrr, offsp_rrrr))
}


# test the function
print(simulate_one_gen_tetraploid(N_RRRR_new = 0, N_RRRr_new = 0, N_RRrr_new = 0, N_Rrrr_new = 0, N_rrrr_new = 1000, 0.01, sel_coeffs = c(0,0.1,0.2,0.1,0), 0.01))
```

# simulating population tetraploid

```{r}
# Function to simulate multiple generations for tetraploid genotypes
simulate_pop_tetraploid <- function(N_init_RRRR, N_init_RRRr, N_init_RRrr, N_init_Rrrr, N_init_rrrr, decay_rate, sel_coeffs, mut_rate, t_max) {
  
  # Create a matrix to save population results
  pop_matrix <- matrix(NA, nrow = t_max + 1, ncol = 5)
  colnames(pop_matrix) <- c("RRRR", "RRRr", "RRrr", "Rrrr", "rrrr")
  pop_matrix[1,] <- c(N_init_RRRR, N_init_RRRr, N_init_RRrr, N_init_Rrrr, N_init_rrrr)
  
  # Initialize the population for the first generation
  pop_current <- c(N_init_RRRR, N_init_RRRr, N_init_RRrr, N_init_Rrrr, N_init_rrrr)
  
  # Run the simulation for t_max generations
  for (t in 1:t_max) {
    pop_next <- simulate_one_gen_tetraploid(pop_current[1], pop_current[2], pop_current[3], pop_current[4], pop_current[5], decay_rate, sel_coeffs, mut_rate)
    pop_matrix[t + 1,] <- pop_next
    pop_current <- pop_next
    
    # Early stop condition: if population drops to zero or exceeds a threshold
    if (sum(pop_next) == 0 || sum(pop_next) >= 1000 * sum(pop_matrix[1,])) {
      break
    }
  }
  
  return(pop_matrix)
}


# Run the simulation and store the output
pop_result <- simulate_pop_tetraploid(N_init_RRRR = 0, N_init_RRRr = 0, N_init_RRrr = 0, N_init_Rrrr = 0, N_init_rrrr = 1000, decay_rate = 0.17, sel_coeffs = c(0,0.165,0.5,0.164,0), mut_rate = 0.01, t_max = 50)

print(pop_result)


# Calculate finite population
#finite_population <- rowSums(pop_result, na.rm = TRUE)

# Check for NA values and stop plotting if found
#if (any(is.na(finite_population))) {
#  stop("Plot creation stopped: Data contains NA values.")
#} else {
#  plot(x_range, finite_population, type='l', ylim=c(0, max(finite_population, na.rm = TRUE)), xlab="Generation", ylab="Population Size", col="black")
#}

#adding lines for the genotype values
#lines(x_range, pop_result[,1], col="blue", lty=1)  # RRRR
#lines(x_range, pop_result[,2], col="red", lty=2)   # RRRr
#lines(x_range, pop_result[,3], col="green", lty=3) # RRrr
#lines(x_range, pop_result[,4], col="purple", lty=4)# Rrrr
#lines(x_range, pop_result[,5], col="orange", lty=5)# rrrr

#legend("topright", legend=c("RRRR", "RRRr", "RRrr", "Rrrr", "rrrr"), col=c("blue", "red", "green", "purple", "orange"), lty=1:5)
```

# selection coefficient table creation (heterozygote advantage)

Here we are generating a selection coefficiency data table to use for
our analysis. Our decay rate is going to be 0.175, so we start the
highest value a little bit lower than the decay rate.

```{r}
create_sel_coeffs <- function(s_number, biggest_S, inc) {
  
 sel_coeffs_table <- matrix(0, nrow = s_number + 1, ncol = 5)
  colnames(sel_coeffs_table) <- c("sc_low", "sc_medium", "sc_high", "sc_medium", "sc_low")
  s <- 0.15
  sc_current <- c(0,s/2, s ,s/2, 0)
  # Run the table creation
  for (i in 1:s_number + 1) {
    sc_next <- c(0,s/2, s ,s/2, 0)
    sel_coeffs_table[i,] <- sc_next
    sc_current <- sc_next
    
    #growing s by increment(inc) we set until it's too big
    if (s <= biggest_S) {
      s <- (s + inc)
      }
   }
  return(sel_coeffs_table)
   }

print(create_sel_coeffs(s_number = 20, biggest_S = 3, inc = 0.05))

#create the table to use later
seltable <- create_sel_coeffs(s_number = 20, biggest_S = 3, inc = 0.05)
print(seltable[3,])

#the down you go in the table the more heterozygote advantage you have.
```

```{r}
# set some parameters to fixed values
init_RRRR <- 0
init_RRRr <- 0
init_RRrr <- 0
init_Rrrr <- 0
init_rrrr <- 250
mut_rate <- 0.01
max_gen <- 100
decay_rate <- 0.155

#set seed for reproducibility
set.seed(123)

# determine how often to run the simulation for each set of parameters
replicates <- 500

#call the seltable here.
seltable <- create_sel_coeffs(s_number = 42, biggest_S = 3, inc = 0.005)

print(seltable)

# initialize vector - where to collect the probablitys
pr <- numeric(nrow(seltable))
# run the simulation across all chosen parameters
# loop over sel coefs tetraploid
for(i in 1:nrow(seltable)){
  # loop over selection coefficients
    sel_coeffs <- seltable[i, ]
    j<-1
    success <- 0
    repeat {
      j <- j+1
      # increase counter by one
    
      # run the simulation once
      first_run <- simulate_pop_tetraploid(init_RRRR, init_RRRr, init_RRrr, init_Rrrr, init_rrrr, decay_rate, sel_coeffs, mut_rate, max_gen)
      
      #complete cases to not take N/As we generated to fill the matrix
      first_run_with_numbers <- first_run[complete.cases(first_run),]

      #calculating total population size in the end
    total_pop_in_the_end <- sum(as.numeric(first_run_with_numbers[nrow(first_run_with_numbers),]))
    
       # determine if the population survived
      if(total_pop_in_the_end != 0) {
        success <- success + 1
      }
        pr[i] = success/replicates
        
        if(j>=replicates) break
      
    }
}

print(pr)


init_RR = 0
init_Rr = 0
init_rr = 250

#this next line is to try with double mutation rate for diploids,take it out of the comment line for running.
#mut_rate <- (mut_rate * 2)

pr_diploid <- numeric(nrow(seltable))
replicates_diploid <- 500
for(k in 1:nrow(seltable)){
  # loop over selection coefficients
    sel_coeffs <- seltable[k, ]
    l<-1
    success_diploid <- 0
    repeat {
      l <- l+1
      # increase counter by one
      # run the simulation once
      first_run_diploid <- simulate_pop_diploid(init_RR, init_Rr, init_rr, decay_rate, sel_coeffs, mut_rate, max_gen)
      
      #complete cases to not take N/As we generated to fill the matrix
      first_run_with_numbers_diploid <- first_run_diploid[complete.cases(first_run_diploid),]

      #calculating total population size in the end
    total_pop_in_the_end_diploid <- sum(as.numeric(first_run_with_numbers_diploid[nrow(first_run_with_numbers_diploid),]))
    
       # determine if the population survived
      if(total_pop_in_the_end_diploid != 0) {
        success_diploid <- success_diploid + 1
      }
        pr_diploid[k] = success_diploid/replicates_diploid
        if(l>=replicates_diploid) break
      
    }
}


plot(pr * 100, type = "b", col = "purple", pch = 15, xlab = "Heterozygote Advantage",
     ylab = "Survival Rate(%)", main = "Tetraploid", ylim = c(0, 100), las = 1)

plot(pr_diploid * 100, type = "b", col = "orange", pch = 15, xlab = "Heterozygote Advantage",
     ylab = "Survival Rate(%)", main = "Diploid", ylim = c(0, 100), las = 1)

# Plot the survival rate (pr) on the y-axis against heterozygote advantage on the x-axis
plot(pr * 100, type = "b", col = "purple", pch = 15, xlab = "Heterozygote Advantage",
     ylab = "Survival Rate (%)", main = "Tetraploid vs Diploid", ylim = c(0, 100), xaxt = 'n', las = 1)

# Add diploid survival rate to the same plot
lines(pr_diploid * 100, type = "b", col = "orange", pch = 16, lty = 2)

# Custom x-axis labels using values from seltable 3rd column, rotated 90 degrees
x_values <- seltable[, 3] # Use values from the 3rd column of seltable
axis(1, at = 1:length(x_values), labels = x_values, las = 2) # Rotate labels 90 degrees

# Custom y-axis with specific ticks
#y_ticks <- c(25, 50, 90) # Define desired tick marks
#axis(2, at = y_ticks, labels = y_ticks, las = 1) # Add custom y-axis with labels

# Add gridlines for better readability
grid()

# Add legend
legend("bottomright", legend=c("Tetraploid", "Diploid"), col=c("purple", "orange"), pch=c(15, 16), lty=c(1, 2))


```

```{r}
#trying what happens with not heterozygot advantage but the count of R alleles
create_sel_coeffs_R <- function(s_number, biggest_S, inc) {
  
 sel_coeffs_table_R <- matrix(0, nrow = s_number + 1, ncol = 5)
  colnames(sel_coeffs_table_R) <- c("sc_low", "sc_medium", "sc_intermediate", "sc_high", "sc_insane")
  s <- 0.1
  sc_current <- c(0,s/4, s/2 ,(s/4)*3, s)
  # Run the table creation
  for (i in 1:s_number + 1) {
    sc_next <- c(0,s/4, s/2 ,(s/4)*3, s)
    sel_coeffs_table_R[i,] <- sc_next
    sc_current <- sc_next
    
    #growing s by increment(inc) we set until it's too big
    if (s <= biggest_S) {
      s <- (s + inc)
      }
   }
  return(sel_coeffs_table_R)
   }

print(create_sel_coeffs_R(s_number = 20, biggest_S = 3, inc = 0.05))

#create the table to use later
seltable_R <- create_sel_coeffs_R(s_number = 20, biggest_S = 3, inc = 0.05)

#the down you go in the table the more R advantage you have.

```

```{r}

# set some parameters to fixed values
init_RRRR <- 0
init_RRRr <- 0
init_RRrr <- 0
init_Rrrr <- 0
init_rrrr <- 250
mut_rate <- 0.01
max_gen <- 100
decay_rate <- 0.155
# determine how often to run the simulation for each set of parameters
replicates <- 500

#call the seltable here.
seltable_R <- create_sel_coeffs_R(s_number = 42, biggest_S = 3, inc = 0.002)

print(seltable_R)

# initialize vector - where to collect the probablitys
pr <- numeric(nrow(seltable_R))
# run the simulation across all chosen parameters
# loop over sel coefs tetraploid
for(i in 1:nrow(seltable_R)){
  # loop over selection coefficients
    sel_coeffs <- seltable_R[i, ]
    j<-1
    success <- 0
    repeat {
      j <- j+1
      # increase counter by one
    
      # run the simulation once
      first_run <- simulate_pop_tetraploid(init_RRRR, init_RRRr, init_RRrr, init_Rrrr, init_rrrr, decay_rate, sel_coeffs, mut_rate, max_gen)
      
      #complete cases to not take N/As we generated to fill the matrix
      first_run_with_numbers <- first_run[complete.cases(first_run),]

      #calculating total population size in the end
    total_pop_in_the_end <- sum(as.numeric(first_run_with_numbers[nrow(first_run_with_numbers),]))
    
       # determine if the population survived
      if(total_pop_in_the_end != 0) {
        success <- success + 1
      }
        pr[i] = success/replicates
        
        if(j>=replicates) break
      
    }
}

print(pr)


init_RR = 0
init_Rr = 0
init_rr = 250

#this next line is to try with double mutation rate for diploids,take it out of the comment line for running.
#mut_rate <- (mut_rate * 2)

pr_diploid <- numeric(nrow(seltable_R))
replicates_diploid <- 500
for(k in 1:nrow(seltable_R)){
  # loop over selection coefficients
    sel_coeffs <- seltable_R[k, ]
    l<-1
    success_diploid <- 0
    repeat {
      l <- l+1
      # increase counter by one
      # run the simulation once
      first_run_diploid <- simulate_pop_diploid(init_RR, init_Rr, init_rr, decay_rate, sel_coeffs, mut_rate, max_gen)
      
      #complete cases to not take N/As we generated to fill the matrix
      first_run_with_numbers_diploid <- first_run_diploid[complete.cases(first_run_diploid),]

      #calculating total population size in the end
    total_pop_in_the_end_diploid <- sum(as.numeric(first_run_with_numbers_diploid[nrow(first_run_with_numbers_diploid),]))
    
       # determine if the population survived
      if(total_pop_in_the_end_diploid != 0) {
        success_diploid <- success_diploid + 1
      }
        pr_diploid[k] = success_diploid/replicates_diploid
        if(l>=replicates_diploid) break
      
    }
}


plot(pr * 100, type = "b", col = "purple", pch = 15, xlab = "R Advantage",
     ylab = "Survival Rate(%)", main = "Tetraploid", ylim = c(0, 100), las = 1)

plot(pr_diploid * 100, type = "b", col = "orange", pch = 15, xlab = "R Advantage",
     ylab = "Survival Rate(%)", main = "Diploid", ylim = c(0, 100), las = 1)

# Plot the success rate (pr) against R advantage
plot(pr * 100, type = "b", col = "purple", pch = 15, xlab = "R Advantage",

ylab = "Survival Rate (%)", main = "Tetraploid vs Diploid", ylim = c(0, 100), xaxt = 'n', las = 1)

# Add diploid survival rate to the same plot
lines(pr_diploid * 100, type = "b", col = "orange", pch = 16, lty = 2)

# Custom x-axis labels using values from seltable_R 5th column, rotated 90 degrees
x_values <- seltable_R[, 5] # Use values from the 5th column of seltable_R
axis(1, at = 1:length(x_values), labels = x_values, las = 2) # Rotate labels 90 degrees

# Add gridlines for better readability
grid()

# Add legend
legend("bottomright", legend=c("Tetraploid", "Diploid"), col=c("purple", "orange"), pch=c(15, 16), lty=c(1, 2))

```
